Learning Fast Quadruped Robot Gaits with the RL PoWER Spline Parameterization

### 2012-06-18 paper_aimsa/experimentalSetup.tex
Next round of edits.

\figp{smartCrop}{0.7}{The red line is the smart cropping border, $g = j1 +  j2$. The dashed grey lines are the physical limits of the joints. As shown in \figref{sixLegPositions}, the fierce movements are prevented by mapping these out-of-range positions to the border line as shown in the graph. The value of $g$ cannot drop below 730 otherwise $j1$ and $j2$ are projected right to the red line.  }

\figp{sixLegPositions}{1}{The left two positions are allowed; middle
  two positions are on the red border line from \figref{smartCrop};
  right two positions are deemed as ''extremely fierce'', thus are
  disallowed. Future work may involve studying a larger range of $g$ for increased performance.}

>>>>

\editbox{Much better, but I would add an intuitive explanation of what the
  smart cropping accomplishes. Old method: crop each servo
  individually (square). New: individual + triangular chop to prevent
  legs going under robot.}

\figp{smartCrop}{0.7}{The red line is the smart cropping border, $g = j1 +  j2$. The dashed grey lines are the physical limits of the joints. As shown in \figref{sixLegPositions}, the fierce movements are prevented by mapping these out-of-range positions to the border line as shown in the graph. The value of $g$ cannot drop below 730 otherwise $j1$ and $j2$ are projected perpendicularly to the red line.  }

\figp{sixLegPositions}{1}{The left two positions are allowed; middle
  two positions are on the red border line from \figref{smartCrop};
  right two positions are deemed as ``extremely fierce'', thus are
  disallowed. Future work may involve studying a larger range of $g$ for increased performance.}

### 2012-06-18 paper_aimsa/relatedWork.tex
Next round of edits.

Many attempts have been made on robot gaits learning using machine
learning algorithms, producing competitive results \edit{Change these
  all to citations} (Chernova and Veloso, 2005; Hornby et al., 2005;
Zykov et al., 2004; Clune et al., 2009a, 2011, 2009b,c; T´ellez et
al., 2006; Valsalam and Miikkulainen, 2008). In fact, Jason et
al. tested and compared six different algorithms for quadruped robot
gait learning in their studies. HyperNEAT, a generative encoding
neural network algorithm, achieved the best performance among all six
methods. In Kormushev et al. (2011)'s bipedal robot energy reduction
research, a reinforcement learning method was used to optimize the
performance. The results from Kormushev's research showed this
algorithm's potential for walking problems. This reinforcement
learning method still needs more tests for further explorations.

>>>>

Many attempts have been made on robot gaits learning using machine
learning algorithms, producing competitive results \edit{Change these
  all to citations} (Chernova and Veloso, 2005; Hornby et al., 2005;
Zykov et al., 2004; Clune et al., 2009a, 2011, 2009b,c; T´ellez et
al., 2006; Valsalam and Miikkulainen, 2008). In fact, Jason et
al. tested and compared six different algorithms for quadruped robot
gait learning in their studies. HyperNEAT, a generative encoding
neural network algorithm, achieved the best performance among all six
methods. In previous bipedal robot energy reduction
research \citep{kormushev2011bipedal-walking-energy}, a reinforcement learning method was used to optimize the
performance. The results from this research showed the
algorithm's potential for walking problems. This reinforcement
learning method still needs more tests for further explorations.

### 2012-06-18 paper_aimsa/results.tex
Next round of edits.

\section{Results}
The results for the gaits evolved by RL PoWER are shown in \figref{rlPerform} and \tabref{results}.The HyperNEAT performance is shown in \figref{hyperPerform}. A total of 900 evaluations were performed for RL PoWER (300 for each of three runs). The reason these two algorithms have different number of trials is that runs continued until the performance plateaued, which we defined as when there was no improvement during the last third of a run. Overall, the RL PoWER gaits were
faster by far, beating the HyperNEAT when comparing
either average or best gaits.

@@

\begin{table}
\begin{center}
\begin{tabular}{|r|c|c|c||c|}
\hline
                                & Average & Best Single & Std. Dev \\
\hline                                    
\hline                                    
RL PoWER                       & {\bf 7.62}  & 11.05  &    2.1   \\
\hline
HyperNEAT                      & 6.28   & 9.7 & 1.26   \\
\hline
\end{tabular}
\caption{The average and standard deviation of the best gaits found
  for both algorithm during each of three runs, in cm/sec.}
\tablabel{results}
\end{center}
\end{table}

We believe that this is because RL PoWER uses evolvable splines to represent gaits As explained earlier,  
The single best gait found during this study had a speed of 11.09 cm/sec,  16.3\% better than the best HyperNEAT gait. \figref{rlPattern} shows a typical RL PoWER gait that had high fitness. Compared with \figref{neat_110115_211410_00000_002_filt_zoom}, the pattern of RL PoWER's motion is less
complex but more regular than the HyperNEAT, but producing higher speed. As shown in \figref{rlPattern}
, multiple motors are more coordinated than the HyperNEAT gait.

>>>>

\section{Results}

The results for the gaits evolved by RL PoWER are shown in \figref{rlPerform} and \tabref{results}.The HyperNEAT performance is shown in \figref{hyperPerform}. A total of 900 evaluations were performed for RL PoWER (300 for each of three runs). The reason these two algorithms have different number of trials is that runs continued until the performance plateaued, which we defined as when there was no improvement during the last third of a run. Overall, the RL PoWER gaits were
faster by far, beating the HyperNEAT when comparing
either average or best gaits.

@@

\begin{table}
\begin{center}
\begin{tabular}{|r|c|c|c|}
\hline
                                & Best Single & Average  & Std. Dev \\
\hline                                    
\hline                                    
RL PoWER                       & {\bf 11.05} & {\bf 7.62}   &    2.1   \\
\hline
HyperNEAT                       & 9.7 & 6.28   & 1.26   \\
\hline
\end{tabular}
\caption{The average and standard deviation of the best gaits found
  for both algorithm during each of three runs, in cm/sec.}
\tablabel{results}
\end{center}
\end{table}

%\edit{Add Kyrre's results to table (one line for tested on his robot (better) and one for tested on our robot (not as good).}

We believe that this is because RL PoWER uses evolvable splines to represent gaits As explained earlier,  
The single best gait found during this study had a speed of 11.09 cm/sec \edit{11.05 or 11.09? See table 2},  16.3\% better than the best HyperNEAT gait. \figref{rlPattern} shows a typical RL PoWER gait that had high fitness. Compared with \figref{neat_110115_211410_00000_002_filt_zoom}, the pattern of RL PoWER's motion is less
complex but more regular than the HyperNEAT, but producing higher speed. As shown in \figref{rlPattern}
, multiple motors are more coordinated than the HyperNEAT gait.
